---
title: "Time Series Analysis"
author: "Yeh"
date: "2019年6月7日"
output: 
  html_document: 
     theme: united
     toc : true
     number_sections: true
     toc_depth : 4
     toc_float : 
      collapsed : true
      smooth_scroll : true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(TSA)
library(forecast)
library(tseries)


# step 1:
## please set working directory and file name
setwd('/Users/yeyuting/Desktop/NCCU/時間數列分析/HW3/')
rawdata = read.csv('機動車輛及道路交通事故分析.csv')


# step 2:
## choose some columns worth to be analysed.
potential_columns = c("肇事總件數.件.", "A1類件數.件.", "A2類件數.件.")
potential_data = as.data.frame(rawdata[,potential_columns])
## Rename to English, Chinese words always exist some problems.
colnames(potential_data) = c('Total', 'A1', 'A2')


# step:3
## target column, we will analysis this data
## You can choose target column after checking output of EDA
target_column = 'A2'
target_data = potential_data[,target_column]

# step 4:
## How many data you would set to be predict for testing
forecast_num = 14 ## from 2018/1 to 2019/2, total 14 months

```


# WorkFlow

0. Exploratory Analysis:
 a. ts-plot : overview whole data (trend or outliers)
 b. box-plot by month : monthly effect

1. De-Trend : 
 a. Determinist Trend(linear model) 
 b. Stochastic Trend (Difference)
 
2. De-Seasonal : 
 a. Determinist Seasonal(season as dummy variable) 
 b. Stochastic seasonal(Difference with lag)
 
3. Fit Model ARMA(p, q) to residuals

4. Diagnose model(last residuals should be White Noise)

5. Simulate prediction 




# Exploratory Analysis

## Overview Potential Data

```{r}
DT::datatable(potential_data)

summary(potential_data)


potential_data_ts = ts(potential_data,
                       start = c(2000,1),
                       freq = 12)
plot(potential_data_ts, main = 'ts-plot')
boxplot(potential_data_ts)

```

Due to extremely low numbers of A1 class and the similarity between Total and A2 class, let's focus on A2 class as our research target.

## Target Data Summary

```{r}

target_data_ts = ts(target_data,start = c(2000,1),freq = 12)


boxplot(target_data_ts ~cycle(target_data_ts ),
        main = target_column , 
        xlab = 'month',
        ylab = target_column )

```

## Training Data
```{r}
len = length(target_data)
data = target_data[1:(len-forecast_num)]
data_ts = ts(data,start = c(2000,1),freq = 12)

test = target_data[(len-forecast_num+1):len]
test_ts = ts(test,end = c(2019,2),freq = 12)

```


# Determinist Method

## Determinist-trend

### Linear Model with intercept

```{r}
model_lm = lm(data_ts~time(data_ts))
summary(model_lm)
```


```{r}
plot(data_ts, main = 'Linear model with intercept');abline(reg = model_lm,col = 'red')

```

### Residual Analysis

```{r}
## original : for follow-up analysis
res_lm = residuals(model_lm)
plot(y=res_lm ,x=as.vector(time(data_ts)),
xlab='Time',ylab='Residuals',type='o',
main = 'Residuals : LM de-trend')

## standard : observe outliers (some issues sor events)
std_res_lm = ts(rstudent(model_lm),
                start = c(2000,1),freq = 12)
plot(y=std_res_lm ,x=as.vector(time(std_res_lm)),
xlab='Time',ylab='Standardized Residuals',type='o',
main = 'Satndard residuals : LM de-trend');abline(a = 3,b = 0, col = 'red', lty = 2);abline(a = -3,b = 0, col = 'red', lty = 2);abline(a = 2,b = 0, col = 'orange', lty = 2);abline(a = -2,b = 0, col = 'orange', lty = 2);abline(a = 0,b = 0, col = 'grey', lty = 2)

```



```{r}
hist(std_res_lm ,
     breaks = 30,
     xlab='Standardized Residuals', 
     main = 'Residuals Histogram')

qqnorm(std_res_lm , main = 'QQ plot');qqline(std_res_lm , col = 'red')

ks.test(std_res_lm ,"pnorm")

acf(res_lm, main = 'ACF of Residuals', lag.max = 60)
```

After de-trend process, residuals are considered to be **normal distribution** , but **not independent** random variable. Therefore, it's clearly proper to treat the data as a time series.  



### Linear Model w.o. intercept
```{r}
model_lm_woi = lm(data_ts~time(data_ts)-1)
summary(model_lm_woi)

plot(data_ts, main='Linear Model w.o. intercept');abline(reg = model_lm_woi, col = 'red')


plot(data_ts, main='Linear Model w.o. intercept (start at (0,0))',
     xlim = c(0,2018), ylim = c(0,26000));abline(reg = model_lm_woi, col = 'red')

```

According to the plot above, the linear *model without intercept may not proper* to fit this data due to the restriction of passing through origin coordinates.


## Determinist-season

Attention : We continue the seasonal analysis with residual data from de-trend process instead of original data.


### Summary

```{r}
season_data = res_lm

season_data_ts = ts(season_data,start = c(2000,1),freq = 12)

plot(y=season_data_ts  ,x=as.vector(time(season_data_ts )),
xlab='Time',ylab='Residuals',type='o',
main = paste0('Residuals of ',target_column,' after de-trend'))

boxplot(season_data_ts~cycle(season_data_ts),
        main = paste0('Residual of ',
                      target_column , 
                      ' after de-trend'), 
        xlab = 'month',
        ylab = 'Residual')
```

It's obviously that the occurrence of motor vehicle accident in **February is lower than other months**. Since means of twelves months are not equal,the seasonal effect should'nt be ignored.





### Seasonal Model w.o intercept

```{r}
mon=season(season_data_ts) 

model_season = lm(season_data_ts~mon-1)

summary(model_season)


model_season_fit = ts(fitted(model_season),start = c(2000,1),freq = 12)

plot(model_season_fit , 
     main = 'Seasonal model w.o. intercept',
     col = 'red',
     ylim = c(min(season_data_ts),max(season_data_ts)));points(season_data_ts, col = 'black')


## residual
res_season = residuals(model_season)
res_season_ts = ts(res_season, 
                      start = c(2000,1),
                      freq = 12)


plot(y=res_season_ts ,x=as.vector(time(res_season_ts)),
xlab='Time',ylab='Residuals',type='o',
main = 'Residuals : de-trend & de-season')

### standard residual
std_res_season = rstudent(model_season)

plot(y=std_res_season  ,x=as.vector(time(std_res_season)),
xlab='Time',ylab='Standardized Residuals',type='o',
main = 'Satndard Residuals : de-trend & de-season');abline(a = 3,b = 0, col = 'red', lty = 2);abline(a = -3,b = 0, col = 'red', lty = 2);abline(a = 2,b = 0, col = 'orange', lty = 2);abline(a = -2,b = 0, col = 'orange', lty = 2);abline(a = 0,b = 0, col = 'grey', lty = 2)

hist(std_res_season ,
     breaks = 30,
     xlab='Standardized Residuals', 
     main = 'Residuals Histogram')

qqnorm(std_res_season , main = 'QQ plot');qqline(std_res_lm , col = 'red')

ks.test(std_res_season ,"pnorm")

```

According to the report, **Jan, Feb, Apr and Dec are significant months** in seasonal model; On the other side ,the plot of fitted values showes that the seasonal model still doesn't fully explain the data.

After de-trend and de-season process, now lets *check the acf and pacf and try to fit ARMA model*;
By the way, interestingly, ks-test still doesn't reject the hypothesis of normal-distribution assumption.


### Seasonal Model with intercept

```{r}
month_=season(season_data_ts) 

model_season_wi=lm(season_data_ts~month_)

summary(model_season_wi)


model_season_wi_fit = ts(fitted(model_season_wi),start = c(2000,1),freq = 12)

ts.plot(model_season_fit,model_season_wi_fit,col = c(2,4),lwd = c(2,2),lty = c(2,3), main = 'Seasonal Means Model',
     ylim = c(min(season_data_ts),max(season_data_ts)));points(season_data_ts, col = 'black')

```

Otherwise, it excludes the January effect in *seasonal model with intercept*, but fitted values are totally the same with ones without intercept; for the sake, the following analysis output will not differ with each seasonal model result.


### Alternative : Harnonic Function

```{r}
har =harmonic(season_data_ts, 6)

model_har = lm(season_data_ts~har)

summary(model_har)


model_har_fit = ts(fitted(model_har),
                   start = c(2000,1),freq = 12)

ts.plot(season_data_ts, lty = c(1,3), col=c(1,4),
        main = 'Harmonic(6)', 
        ylim = c(min(model_har_fit,season_data_ts),
                max(model_har_fit,season_data_ts)));points(season_data_ts);lines(model_har_fit,col="red")

```




## ARMA

### ACF and PACF
```{r}
arma_data = res_season
arma_data_ts = ts(arma_data,start = c(2000,1),freq = 12)

plot(y=arma_data_ts  ,x=as.vector(time(arma_data_ts)),
xlab='Time',ylab='Residuals',type='o',
main = paste0('Residuals of ',
              target_column ,
              ' after de-trend and de-season'))



acf(arma_data, lag = 60, main = 'ACF of Residuals after de-trend and de season')
pacf(arma_data, lag = 60, main = 'PACF of Residuals after de-trend and de season')
```

The tail-off ACF plot with cuts-off PACF plot implies the AR(p) model. Because of cuts-off after lag 1 in PACF plot, we first give AR(1) or SARMA(1,0)(1,0)[6] a try.

### AR(1)

#### Model
```{r}
model_ar1 = Arima(arma_data_ts, order = c(1,0,0))
summary(model_ar1)

model_ar1_fit = fitted(model_ar1)

ts.plot(arma_data_ts, model_ar1_fit,col = c(1,2),lwd = c(2,1),lty = c(2,1), main = 'AR(1) Model',
     ylim = c(min(arma_data_ts),max(arma_data_ts)));points(arma_data_ts, col = 'black')

```




#### Residuals
```{r}
## residual
res_ar1  = residuals(model_ar1) ## model_ar1$residuals
res_ar1_ts = ts(res_ar1, start = c(2000,1), freq = 12)


plot(y=res_ar1_ts ,x=as.vector(time(res_ar1_ts)),
xlab='Time',ylab='Residuals',type='o',
main = 'Residuals of AR(1) after de-trend & de-season')

## standard residual
## rstudent not work to Arima object, but I found rstandard's output not equal to rstudent with seasonal model

std_res_ar1 = rstandard(model_ar1) 

plot(y=std_res_ar1  ,x=as.vector(time(std_res_ar1)),
xlab='Time',ylab='Standardized Residuals',type='o',
main = 'Satndard Residuals of AR(1) after de-trend & de-season');abline(a = 3,b = 0, col = 'red', lty = 2);abline(a = -3,b = 0, col = 'red', lty = 2);abline(a = 2,b = 0, col = 'orange', lty = 2);abline(a = -2,b = 0, col = 'orange', lty = 2);abline(a = 0,b = 0, col = 'grey', lty = 2)


hist(std_res_ar1 ,
     breaks = 30,
     xlab='Standardized Residuals', 
     main = 'Residuals Histogram')

qqnorm(std_res_ar1 , main = 'QQ plot');qqline(std_res_ar1 , col = 'red')

ks.test(std_res_ar1 ,"pnorm")

```

#### White Noise Diagnose
```{r}
acf(as.vector(res_ar1), lag = 60, main = 'ACF : Residuals of AR(1)')

pacf(as.vector(res_ar1), lag = 60, main = 'PACF : Residuals of AR(1)')
```


### SARIMA(1,0,0)(1,0,0)[6]

#### Model
```{r}
model_ar1010 = Arima(arma_data_ts, 
                  order = c(1,0,0),
                  seasonal = list(order = c(1,0,0),
                                  period = 6))
summary(model_ar1010)

model_ar1010_fit = fitted(model_ar1010)

ts.plot(arma_data_ts, model_ar1010_fit,col = c(1,2),lwd = c(2,1),lty = c(2,1), main = 'SARIMA(1,0,0)(1,0,0)[6] Model',
     ylim = c(min(arma_data_ts),max(arma_data_ts)));points(arma_data_ts, col = 'black')

```

#### Residuals
```{r}
## residual
res_ar1010  = residuals(model_ar1010) ## model_ar1$residuals
res_ar1010_ts = ts(res_ar1010, start = c(2000,1), freq = 12)


plot(y=res_ar1010_ts ,x=as.vector(time(res_ar1010_ts)),
xlab='Time',ylab='Residuals',type='o',
main = 'Residuals of SARIMA(1,0,0)(1,0,0)[6]')

## standard residual
## rstudent not work to Arima object, but I found rstandard's output not equal to rstudent with seasonal model

std_res_ar1010 = rstandard(model_ar1010) 

plot(y=std_res_ar1010  ,x=as.vector(time(std_res_ar1010)),
xlab='Time',ylab='Standardized Residuals',type='o',
main = 'Satndard Residuals of SARIMA(1,0,0)(1,0,0)[6]');abline(a = 3,b = 0, col = 'red', lty = 2);abline(a = -3,b = 0, col = 'red', lty = 2);abline(a = 2,b = 0, col = 'orange', lty = 2);abline(a = -2,b = 0, col = 'orange', lty = 2);abline(a = 0,b = 0, col = 'grey', lty = 2)


hist(std_res_ar1010 ,
     breaks = 30,
     xlab='Standardized Residuals', 
     main = 'Residuals Histogram')

qqnorm(std_res_ar1010 , main = 'QQ plot');qqline(std_res_ar1010 , col = 'red')

ks.test(std_res_ar1010 ,"pnorm")

```

#### White Noise Diagnose
```{r}
acf(as.vector(res_ar1010), lag = 60, main = 'ACF : Residuals of SARIMA(1,0,0)(1,0,0)[6]')

pacf(as.vector(res_ar1010), lag = 60, main = 'PACF : Residuals of SARIMA(1,0,0)(1,0,0)[6]')
```



### Potential parameters for ARMA(p,q)

#### Method I : EACF
```{r}
eacf(arma_data)

# eacf_info = eacf(arma_data)
# eacf_sig = eacf_info$eacf
# eacf_sig < 0.05



```

<img src = 'https://imgur.com/pURnTsc.png'></img>




By EACF matrix, ARMA(1,1) is our candidate for modeling.

#### Method II : auto.arima
Attention: Because we early import deterministic trend and seasonal model; to prevent it from stochastic effect, we set zero to parameters d, D, max.P and max.Q.

```{r}
auto_arima = auto.arima(arma_data_ts,
           d = 0, D = 0,
           trace = T)
auto_parameters = auto_arima$arma #(p,q,P,Q,s,d,D)
```

By auto.arima function in "forecast" package, the best model is SARIMA(1,0,1)(1,0,1)[12] with zero mean 



So, we have three candidate:
a. ARMA(1,1)
b. SARMA(1,0,1)(1,0,1)[12]

### Candidate ARMA(p,q)
```{r}
## ARMA(1,1)
model_arma11 = Arima(arma_data_ts, order = c(1,0,1))
model_arma11_fit = fitted(model_arma11)
res_arma11 = residuals(model_arma11)
std_res_arma11 = rstandard(model_arma11) 

## auto arima
auto_p = auto_parameters[1]
auto_q = auto_parameters[2]
auto_P = auto_parameters[3]
auto_Q = auto_parameters[4]
auto_s = auto_parameters[5]
auto_d = auto_parameters[6]
auto_D = auto_parameters[7]

model_arma_auto = Arima(arma_data_ts, 
                        order = c(auto_p,auto_d,auto_q),
                        seasonal = list(
                          order = c(auto_P,auto_D,auto_Q),
                          period = auto_s)
                        )
model_arma_auto_fit = fitted(model_arma_auto)
res_arma_auto = residuals(model_arma_auto)
std_res_arma_auto = rstandard(model_arma_auto) 

model_form = paste0("ARIMA(",
                    auto_p,",",
                    auto_d,",",
                    auto_q,")(",
                    auto_P,",",
                    auto_D,",",
                    auto_Q,")[",
                    auto_s,"]")
## plot
ts.plot(arma_data_ts, 
        model_arma11_fit,
        model_arma_auto_fit,
        col = c(1,2,4),
        lwd = c(1,2,2),
        lty = c(1,2,3), 
        
        main = paste0('ARMA(1,1) and  ',model_form,'Model'),
        ylim = c(min(arma_data_ts),
              max(arma_data_ts)))
# ;points(
#         arma_data_ts, col = 'black');legend(
#         'top', 
#         legend = c('data','arma11',
#                    paste0(model_form)),
#         col = c(1,2,4),lty = 1,cex = 0.5
#                       )



```


## Diagnose

### Outlier

```{r}
plot(y=std_res_arma11  ,
     x=as.vector(time(std_res_arma11)),
     xlab='Time',
     ylab='Standardized Residuals',
     type='o',
     main = 'Satndard Residuals of ARMA(1,1) after de-trend & de-season');abline(a = 3,b = 0, 
     col = 'red', lty = 2);abline(a = -3,b = 0, 
     col = 'red', lty = 2);abline(a = 2,b = 0, 
     col = 'orange', lty = 2);abline(a = -2,b = 0, 
     col = 'orange', lty = 2);abline(a = 0,b = 0, 
     col = 'grey', lty = 2)


plot(y=std_res_arma_auto,
     x=as.vector(time(std_res_arma_auto)),
     ylim = c(-3.5, 3.5),
     xlab='Time',
     ylab='Standardized Residuals',
     type='o',
     main = paste0('Satndard Residuals of ARMA(',
     auto_p,',',auto_q,
     ') after de-trend & de-season'));abline(a = 3,b = 0, 
     col = 'red', lty = 2);abline(a = -3,b = 0, 
     col = 'red', lty = 2);abline(a = 2,b = 0, 
     col = 'orange', lty = 2);abline(a = -2,b = 0, 
     col = 'orange', lty = 2);abline(a = 0,b = 0, 
     col = 'grey', lty = 2)
```


### White Noise

```{r}
acf(as.vector(res_arma11), lag = 60, main = 'ACF : Residuals of ARMA(1,1)')
pacf(as.vector(res_arma11), lag = 60, main = 'PACF : Residuals of ARMA(1,1)')


acf(as.vector(res_arma_auto), lag = 60, 
    main = paste0('ACF : Residuals of ARMA(',auto_p,',',auto_q,')'))
pacf(as.vector(res_arma_auto), lag = 60, 
     main = paste0('PACF : Residuals of ARMA(',auto_p,',',auto_q,')'))


```

Consider that the residuals had better be whote noise, 
neither ARMA(1,1) nor ARMA(5,2) meets the ideal result because of the peaks with lag 12 on ACF plots, which means it's definitely not independent white noise.Therefore, lets give stochastic method a try.




# Stochastic Method

## Stochastic-Trend 

### Dickey Fuller Test

H0: non-stationary with lag k 
H1: stationary with lag k


```{r}
## stationary
adf.test(data_ts, alternative = c("stationary"),k = 1)
adf.test(data_ts, alternative = c("stationary"),k = 2)
adf.test(data_ts, alternative = c("stationary"),k = 3)
## non stationary
adf.test(data_ts, alternative = c("stationary"),k = 4)
adf.test(data_ts, alternative = c("stationary"),k = 5)
adf.test(data_ts, alternative = c("stationary"),k = 6)
adf.test(data_ts, alternative = c("stationary"),k = 12)

```

Althogh the above shows outputs of adf.test, I don't really need it and won't interpret the test.


### ACF before difference
```{r}
acf(data_ts, lag = 60)
pacf(data_ts, lag = 60)
```
The tail-off acf, nearly equal to 1.0 starting at lag 1, strongly implies the stochastic trend effect. So I try to import non-seasonal part of the arima model, which means lets take difference with lag 1.


### Difference
```{r}
## method I : arima model with d = 1
model_stoch_trend = Arima(data_ts, order = c(0,1,0))
model_stoch_trend_fit = fitted(model_stoch_trend)

ts.plot(data_ts,model_stoch_trend_fit ,
        col = c(1,2),lwd = c(1,2),lty = c(1,2), 
        main = 'Stochastic Trend Model',
     ylim = c(min(data_ts),max(data_ts)))

res_stoch_trend = residuals(model_stoch_trend)

## method II : diff 
diff(data_ts)==residuals(model_stoch_trend)

```

### ACF after difference
```{r}
acf(as.vector(res_stoch_trend), lag = 60)
pacf(as.vector(res_stoch_trend), lag = 60)
```

Compare determinist linear model to stochastic trend model, I think it's better to choose the latter one because I noticed that acf after difference shows repeatedly high auto-correlation with lag 12. It may help me more to interpret the further stochastic seasonal model.


## Stochastic-Season

```{r}
stoch_season_data = res_stoch_trend
## method I : arima model with D=1 and period=12
model_stoch_season = Arima(stoch_season_data,
        seasonal = list(order = c(0,1,0), period = 12))

model_stoch_season_fit = fitted(model_stoch_season)

ts.plot(stoch_season_data,
        model_stoch_season_fit ,
        col = c(1,2),lwd = c(1,1),lty = c(1,2), 
        main = 'Stochastic Season Model',
        ylim = c(min(stoch_season_data),
              max(stoch_season_data)))

res_stoch_season = residuals(model_stoch_season)

## method 2 : diff with lag 12
diff(stoch_season_data,lag = 12)==residuals(model_stoch_season)


```


## ARMA

### ACF and PACF
```{r}
## if you dont import stochastic season effect
# stoch_arma_data = res_stoch_trend 

stoch_arma_data = res_stoch_season
plot(y=stoch_arma_data ,
     x=as.vector(time(stoch_arma_data)),
     xlab='Time',ylab='Residuals',type='o',
main = 'Residuals after de-trend and de-season')



acf(as.vector(stoch_arma_data), lag = 60, 
    main = 'ACF of Residuals after de-trend and de season')
pacf(as.vector(stoch_arma_data), lag = 60, 
     main = 'PACF of Residuals after de-trend and de season')
```

With low and irregular acf and pacf, it too hard for me to determine ARMA parameters p and q by  my poor experience.

### Potential parameters for ARMA(p,q)

#### Method I : EACF

It can only partially explain non-seasonal parameters p,q
```{r}
eacf(stoch_arma_data)
```

ARMA(0,1) may be a proper choice

## Method II: auto.arima
```{r}
auto_arima = auto.arima(stoch_arma_data,
           d = 0, D = 0,
           trace = T)

auto_parameters = auto_arima$arma #(p,q,P,Q,s,d,D)
auto_p = auto_parameters[1]
auto_q = auto_parameters[2]
auto_P = auto_parameters[3]
auto_Q = auto_parameters[4]
auto_s = auto_parameters[5]
auto_d = auto_parameters[6]
auto_D = auto_parameters[7]
```

According to the auto.arima, the best model is SARIMA(p=0,d=0,q=1)(P=0,D=0,Q=1) with lag 12 and without drift.

### Candidate SARIMA(p,d,q)(P,D,Q)[s]
```{r}

model_stoch_arma_auto = Arima(stoch_arma_data, 
                        order = c(auto_p,auto_d,auto_q),
                        seasonal = list(
                          order = c(auto_P,auto_D,auto_Q),
                          period = auto_s)
                        )
model_stoch_arma_auto_fit = (stoch_arma_data - model_stoch_arma_auto$residuals)
#model_stoch_arma_auto_fit == fitted(model_stoch_arma_auto)


res_stoch_arma_auto = residuals(model_stoch_arma_auto)
std_res_stoch_arma_auto = rstandard(model_stoch_arma_auto) 

# res_stoch_arma_auto  == model_stoch_arma_auto$residuals

ts.plot(stoch_arma_data,
        model_stoch_arma_auto_fit ,
        col = c(1,2),lwd = c(1,1),lty = c(1,2), 
        main = 'SARMA Model',
        ylim = c(min(stoch_arma_data),
              max(stoch_arma_data)))




```


## Diagnose 

### Outlier

```{r}
# plot(res_stoch_arma_auto) ## just check
plot(y=std_res_stoch_arma_auto,
     x=as.vector(time(std_res_stoch_arma_auto)),
     xlab='Time',
     ylab='Standardized Residuals',
     type='o',
     main = 'Satndard Residuals of ARMA after de-trend & de-season');abline(a = 3,b = 0, 
     col = 'red', lty = 2);abline(a = -3,b = 0, 
     col = 'red', lty = 2);abline(a = 2,b = 0, 
     col = 'orange', lty = 2);abline(a = -2,b = 0, 
     col = 'orange', lty = 2);abline(a = 0,b = 0, 
     col = 'grey', lty = 2)

```


### White Noise

```{r}
acf(as.vector(res_stoch_arma_auto), 
    lag = 60, main = 'ACF : Residuals of SARMA')
pacf(as.vector(res_stoch_arma_auto), 
     lag = 60, main = 'PACF : Residuals of SARMA')
```


Before prediction, let's summarise the model info.
Firstly, we apply linear model or difference into the de-trend process; secondly, we apply seasonal model or difference with a lag into the de-season process.After trend and seasonal effects are removed, we established an ARMA or a Seasonal ARMA model to fit the data and expect that the residuals could be white noise.

Therefore, the whole fitting model of **Determinist Method** would contain three parts : ARMA, Seasonal Model, Trend Model
$$ (1-\Sigma_{i}^{p}{\phi_i B_i})Y_t = (1-\Sigma_{j}^{q}{\theta_j B^j})e_t + (\Sigma{\gamma_k m_k}) + (\beta_1 t  + \beta_0)$$
$$ B : lag \space operator  $$
$$p,q : non-seasonal \space ARMA \space parameter$$

$$m_i : dummy \space variable \space of \space seasonal\space effect$$

$$\gamma : coefficient \space of \space season\space effect$$

$$\beta : coefficient \space of \space trend\space effect$$

The Whole **Stochastic Model** would be like:
$$ (1-B)^d(1-B^s)^D(1-\Sigma_{i}^{p}{\phi_i B^i})(1-\Sigma_{j}^{P}{\Phi_j B^{js}})Y_t = (1-\Sigma_{k}^{q}{\theta_k B^k})(1-\Sigma_{l}^{Q}{\theta_l B^{ls}})e_t $$

$$ B : lag \space operator  $$
$$p,d,q : nonseasonal \space ARMA \space parameter$$
$$P,D,Q,s : seasonal \space ARMA \space parameter$$


### Simplify Whole Stochastic Model

The whole stochastic model go through once difference for de-trend, once defference with lag 12 for de-season,and fitting a seasonal ARMA(0,0,1)(0,0,1)[12] model.Now we can simplify the whole model as a SARIMA(0,1,1)(0,1,1)[12]

#### Stochastic Method
```{r}
model_sarima = Arima(y = data_ts,
        order = c(0,1,1),
        seasonal = list(order = c(0,1,1), 
                        period = 12))

model_sarima_fit = (data_ts - model_sarima$residuals)
res_sarima =  model_sarima$residuals

## I don't know why the residual not the same
res_sarima == res_stoch_arma_auto
res_sarima - res_stoch_arma_auto

## But the acf and pacf pattern are nearly equal
par(mfrow = c(2,2));acf(as.vector(res_sarima), 
    lag = 60, main = 'ACF : Residuals of SARIMA');pacf(as.vector(res_sarima), 
     lag = 60, main = 'PACF : Residuals of SARIMA');acf(as.vector(res_stoch_arma_auto), 
    lag = 60, main = 'ACF : Residuals of whole split models previously');pacf(as.vector(res_stoch_arma_auto), 
     lag = 60, main = 'PACF : Residuals of whole split models previously')
```


#### Determinist Method

```{r}
dummy =  seasonaldummy(ts(data_ts,f=12))
var_time = time(data_ts)
xreg = model.matrix(~dummy + var_time)

model_det_arima = Arima(y = data_ts,
                     xreg = xreg ,
                     include.mean=FALSE ,
                     order = c(1,0,1),
                     seasonal=list(order=c(1,0,1), period=12)
                    )

model_det_arima_fit = (data_ts - model_det_arima $residuals)
res_det_arima  =  model_det_arima$residuals

```


#### Mix Method I

<font size = 4> 
Determinist de-trend + Stochastic de-season + SARIMA
</font>

```{r}
xreg = time(data_ts)

model_mix_arima = Arima(y = data_ts,
                     xreg = xreg,
                     order = c(1,0,1),
                     seasonal=list(order=c(1,1,1), period=12)
                    )

model_mix_arima_fit = (data_ts - model_mix_arima$residuals)
res_mix_arima  =  model_mix_arima$residuals
```


#### Mix Method II

<font size = 4> 
Stochastic de-trend +  Determinist de-season + SARIMA
I met some Non-Stationary Problems
</font>

```{r}
# dummy =  seasonaldummy(ts(data_ts,f=12))
# xreg = model.matrix(~dummy)
# 
# model_mix_arima = Arima(y =data_ts,
#                      xreg = xreg,
#                      order = c(1,0,1),
#                      include.mean=FALSE ,
#                      seasonal=list(order=c(1,0,1), 
#                                    period=12)
#                     )
# 
# model_mix_arima_fit = (data_ts - model_mix_arima$residuals)
# res_mix_arima  =  model_mix_arima$residuals
```




# Prediction

## Stochastic
```{r}
pred_sarima = forecast(model_sarima, h = forecast_num)

plot(pred_sarima);lines(pred_sarima$fitted,col="red");lines(target_data_ts,col="black")


```

## Determinist
```{r}
test_dummy =  seasonaldummy(ts(test_ts,f=12))
test_var_time = time(test_ts)
test_xreg = model.matrix(~test_dummy + test_var_time)

pred_det_arima = forecast(model_det_arima, 
                          h = forecast_num,
                          xreg = test_xreg)

plot(pred_det_arima);lines(pred_det_arima$fitted,col="red");lines(target_data_ts,col="black")
```


## Mix Model 
```{r}
test_xreg = time(test_ts)

pred_mix_arima = forecast(model_mix_arima, 
                          h = forecast_num,
                          xreg = test_xreg)

plot(pred_mix_arima);lines(pred_mix_arima$fitted,col="red");lines(target_data_ts,col="black")
```




<font size = 5> Thanks for reading !</font>


